---
title: "607-Week 9 Assignment Web API's"
author: "Vijaya Cherukuri"
date: "October 28, 2018"
output: html_document
---

# Assignment
> The New York Times web site provides a rich set of APIs, as described here: http://developer.nytimes.com/docs
You'll need to start by signing up for an API key.
Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe.

# Solution

## Libraries 

```{r}
library(jsonlite)
library(DT)
library(tidyr)
library(dplyr)
```

## Get API Key from New York Times website
```{r}
key <- "&api-key=866c0b4f1a5c4bb4a978ed35735f4ccd"
```

## We took books of best sellers where age range is filter as 10
```{r}
url <- "https://api.nytimes.com/svc/books/v3/lists/best-sellers/history.json?age-group=10"
```

## Use jsonlite function to get API data with API key and query for kids books

```{r}
maindata <- fromJSON(paste0(url, key))
```

## Verify the data and tidy

```{r}
head(maindata)

# Verify Class, length and Names
class(maindata) 

length(maindata)

names(maindata)

# Check number of elements in Lists
lapply(maindata, class)

lapply(maindata, length)
```

## Tidy data and create dataframe

```{r cars}
#Remove first three lists
results <- maindata[-c(1:3)]
names(results)

length(results)
class(results)

results_df <- results[["results"]]

class(results_df)

colnames(results_df)

Books <- tbl_df(results_df)
Books <- select(Books, one_of(c("title", "description", "author", "price", "age_group", "publisher")))
Books <- rename(Books, Title = 'title', Description = 'description', Author = 'author', Price = 'price', Ages = 'age_group', Publisher = 'publisher')
Books

datatable(Books)
```


